---
title: "p8105_hw3_yq2251"
author: "TritonD"
date: "10/12/2019"
output: github_document
---
```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(knitr)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

## count how many aisles there are
```{r}
instacart %>%
  group_by(aisle) %>%
  summarize(aisle_n = n())
```

## find out which aisle is most popular
```{r}
instacart %>%
  group_by(aisle) %>%
  summarize(aisle_n = n())%>%
filter(min_rank(desc(aisle_n)) < 2)

```
There are 134 aisles in total, with "fresh vegetables" as the aisle with most items ordered from (n=150609).



## Make a plot that shows the number of items ordered in each aisle with n>10000
```{r}
instacart %>%
  group_by(aisle) %>%
  summarize(aisle_n = n())%>%
filter(aisle_n > 10000)%>%
ggplot(aes(x=aisle, y=aisle_n)) +
  geom_bar(stat="identity") +
  theme(text = element_text(size=12),
         axis.text.x = element_text(angle=90, vjust=0.5)) +
  labs(
    title = "Ordering by aisle plot",
    x = "Aisle",
    y = "Number of items ordered in each aisle",
    caption = "Data from Instacart datasets"
  )
```

## make a table for 3 most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r}
instacart %>%
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(aisle_n_product = n())%>%
  filter(min_rank(desc(aisle_n_product)) < 4)%>%
  knitr::kable(digits = 0)
```


## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name,order_dow) %>% 
  summarize(mean_time = mean(order_hour_of_day))%>%
   pivot_wider(
    names_from = order_dow,
    values_from = mean_time) %>% 
  rename(Sunday="0", Monday="1", Tuesday="2", Wednesday="3", Thursday="4", Friday="5", Saturday="6")%>%
  knitr::kable(digits = 0)
```


## problem 2

